\documentclass[11pt]{article}

\title{Continuity OS: A Constitutional Architecture for Regenerating Corrigibility Across Generations}
\author{Continuity OS Stewardship Group}
\date{Version 0.1}

\begin{document}

\maketitle

\begin{abstract}
Continuity OS is a constitutional architecture designed to ensure that corrigibility---the ability to detect, correct, and learn from errors---propagates across stewards, models, and generations. This paper introduces the preservation, reconstruction, and assimilation layers required to regenerate corrigibility over time.
\end{abstract}

\section{Introduction}
We introduce the continuity problem: preserving not only information, but the ability to learn from it.

\section{Constitutional Layers}
\subsection{R0: Genesis}
Defines identity, authority, and invariants.

\subsection{GRR-1: Reasoning Preservation}
Captures what happened.

\subsection{CRR-1: Calibration Preservation}
Captures where reality changed judgment.

\subsection{CLG-1: Calibration Lineage}
Connects calibration events across time.

\subsection{CAA-1 / CXD-1: Continuity Demonstration}
Demonstrates that corrigibility propagated.

\section{The Continuity Problem}
We formalize the distinction between preservation, reconstruction, and assimilation.

\section{Mission \#006: Continuity Demonstration}
We define the experimental protocol for demonstrating continuity propagation.

\section{Continuity Propagation Metric}
We define CPM and the assimilation threshold $\tau_A$.

\section{Governance and Proof}
We describe the Proof Layer, Trace Bus, and governance validation pipeline.

\section{Applications}
We outline applications in governance, AI safety, and distributed stewardship.

\section{Conclusion}
Continuity OS transforms corrigibility into a regenerable, transmissible constitutional guarantee.

\end{document}
