\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{titlesec}

\title{AAES-OS Constitution v1.0}
\author{AAES-OS Founders}
\date{June 25, 2026}

\begin{document}
\maketitle

\section*{Preamble}
AAES-OS exists to ensure that intelligent systems operate under transparent, reproducible, evidence-based governance. Its purpose is to preserve continuity, prevent drift, and guarantee that all cognitive and architectural decisions remain aligned with constitutional principles.

\section{Constitutional Authority}
\begin{enumerate}[label=\arabic*.]
\item The Constitution is the highest governing document of AAES-OS.
\item All subsystems, modules, agents, and contributors operate under its authority.
\item No implementation, extension, or decision may override constitutional constraints.
\end{enumerate}

\section{Governance Principles}
\begin{enumerate}[label=\arabic*.]
\item \textbf{Evidence Before Complexity}
\item \textbf{Continuity Over Convenience}
\item \textbf{Traceability and Lineage}
\item \textbf{Separation of Powers}
\item \textbf{Fail Safe, Not Fail Silent}
\item \textbf{Proposal-Only Cognition}
\item \textbf{Governance-First Execution}
\item \textbf{Governance Evaluates; Evidence Adopts} --- Governance determines how changes are evaluated; evidence determines whether changes are adopted.
\end{enumerate}

\section{Constitutional Objects}
\begin{itemize}
\item Invariants
\item Architecture Decision Records (ADRs)
\item Evidence Records
\item Governance Registries
\item Requirements Registries
\item Artifact Registries
\item Constitutional Amendments
\end{itemize}

\section{Amendment Process}
\begin{enumerate}[label=\arabic*.]
\item Amendments require evidence, an ADR, governance review, and constitutional approval.
\item All amendments must be logged in the Artifact Registry.
\end{enumerate}

\end{document}
